---
title: "Project 1"
author: "Allen Crane, Nicholas Cellini, Heber Nielsen, Quincy Roundtree, Chris Graves"
date: "June 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Market Backdrop
###The crowded Microbrew landscape continues to expand with 558 Breweries offering a proliferating selection of 2,410 beers across the domestic US market.
###The distribution of Breweries by state is highly variable with concentrations in Colorado, California, and Michigan - homes of "old guard" national brands and brews.
###While median alcohol content (ABV) is solidly clustered in the 5% to 6% range across markets, bitterness (IBU) is much more variable.

##Meet FrankenBeer
###FrankenBeer was created to address a perceived 'untapped' demand for the crisp taste of high alcohol content beer without the bitterness which accompanies existing product offerings in this segment. 
###Our beer scientists have created a novel brewing technique enabling us to deliver 12% ABV at only 33 IBUs. 
###The ABRAcadabraTM  process (Alcohol Bitterness Reduction & Abatement) has performed well within our dedicated team of full time tasters AND in initial test markets across multiple levels of ABV.
###Our FrankenBeer has distinct qualities that make it appealing customers who enjoy strong beers but dislike bitterness.
###The ABRAcadabraTM  process provides Franken Brewery a low-bitterness competitive advantage which extends across all alcohol content levels.
###The Franken brewery seeks to begin distribution of FrankenBeer in three new states where it can most effectively bring this competitive advantage to bear.

##Initial Market Research
###Problem Statement: Our FrankenBeer has distinct qualities that make it appealing to many demographics of beer drinkers. This indicates that a thorough analysis must be done to target receptive markets.
###Current Condition: Our FrankenBeer is a 12% ABV with a 33 IBU. It has tested well in the Seattle, Washington and Boulder, Colorado markets.
###Target Condition: Expand the FrankenBeer to three new markets.

##Initial Market Research 
###The first phase of our work centered on our analysis of the US microbreweries by state. Of particular interest was the Alcohol and Bitterness in the different beer varieties found in each area. 
###Initial testing has been done in the Seattle, Washington and Boulder, Colorado markets with much success. 
###Colorado already has the highest ABV craft beer in the country--Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale (ABV 12.8%).

##Expansion Market Proposal
###To assist with target market identification, our data scientists (in a flash of brilliance) have created the FrankenScoreTM which is the product of the median ABV by state and the median IBU by state.
###This metric sizes the relative advantage Franken's new proprietary technique might have WV, NM, & FL were identified as having the highest market potential.


##Set working directory and get data files

```{r}
setwd("C:\\Users\\allen\\Documents\\SMU Data Science\\MSDS6306 - Doing Data Science\\Lecture 5")
beers1.df <- read.csv("Beers.csv", header=TRUE, sep = ",")
beers.df <- read.csv("Enhanced_beers.csv", header=TRUE, sep = ",")
breweries.df <- read.csv("Breweries.csv", header=TRUE, sep = ",")
```


##1. How many breweries are present in each state?

```{r}
library(plyr)
library(lattice)
brewery_count <- table(breweries.df$State)
brewery_count
barplot(brewery_count, border = "darkblue", col = "aquamarine", main = "Breweries per State", xlab = "States", ylab="Number of Breweries", ylim = range(pretty(c(0, brewery_count))))
```


##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

#### First create a new variable "Brewery_id" (copy of "Brew_ID") that can merge both files
#### Merge data on Brewery_ID
#### Clean up column names

```{r}
breweries.df$Brewery_id <- breweries.df$Brew_ID
MergeBeerBreweries1 <- merge(x=beers1.df, y=breweries.df, by="Brewery_id", all=TRUE)
names(MergeBeerBreweries1) <- c("Brewery_ID", "Beer_Name", "Beer_ID", "ABV", "IBU", "Style", "Ounces", "Brew_ID", "Brewery_Name", "City", "State")
MergeBeerBreweries <- merge(x=beers.df, y=breweries.df, by="Brewery_id", all=TRUE)
names(MergeBeerBreweries) <- c("Brewery_ID", "Beer_Name", "Beer_ID", "ABV", "IBU", "Style", "Ounces", "Brew_ID", "Brewery_Name", "City", "State")
head(MergeBeerBreweries)
tail(MergeBeerBreweries)
```


##3. Report the number of NA's in each column.
####Note: In the original Beers data set, we found that there were 62 NA values for ABV and 1005 NA values for IBU.
####Our team of beer researchers used additional data sources (Beer Advocate and Untapped) to impute more than 200 missing values, which reduced the number of NAs significantly, down to 55 NA values for ABV and 891 NA values for IBU. We will be using this enhanced data set for the remainder of this project, as it adds many more data values for our analysis than we already had.
####Sources: Nick Hould Github page: https://github.com/nickhould/craft-beers-dataset; http://www.untappd.com; http://www.beeradvocate.com


```{r}
summary(MergeBeerBreweries1)
summary(MergeBeerBreweries)
```


##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

#### First create a median function so we can group by State
#### Median of ABV by State
#### Plot bar chart of ABV by State
#### Plot box plot of ABV by State
#### Median of IBU by State
#### Plot bar chart of IBU by State
#### Plot box plot of IBU by State


```{r}
MedianBeerStat <- function(x) {
  c(median = median(x[which(!is.na(x))]))
}
tapply(MergeBeerBreweries$ABV, MergeBeerBreweries$State, MedianBeerStat)
MedianABV <- tapply(MergeBeerBreweries$ABV, MergeBeerBreweries$State, MedianBeerStat)
barplot(MedianABV, main="Median Alcohol By Volume (ABV) by State", xlab="State", ylab="Median ABV", col="darkblue")
boxplot (ABV ~ State, data = MergeBeerBreweries, col = "goldenrod", main = "Alcoholic Content of Beers", subtitle = "Beers Brewed in Each State and the District of Columbia", xlab = "States in the US", ylab = "Beer Alcohol Content")
tapply(MergeBeerBreweries$IBU, MergeBeerBreweries$State, MedianBeerStat)
MedianIBU <- tapply(MergeBeerBreweries$IBU, MergeBeerBreweries$State, MedianBeerStat)
barplot(MedianIBU, main="Median International Bitterness Units (IBU) by State", xlab="State", ylab="Median IBU", col="darkblue")
boxplot (IBU ~ State, data = MergeBeerBreweries, col = "rosybrown", main = "Bitterness Scores by State", subtitle = "Beers Brewed in Each State and the District of Columbia", xlab = "States in the US", ylab = "Beer Bitterness Content")
```


##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

#### Colorado has the maximum alcoholic beer "Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale" (ABV 12.8%), in terms of individual beers, while Oregon has the most bitter beer "Bitter Bitch Imperial IPA" (IBU 138), in terms of individual beers.
#### Using the median ABV and IBU data respectivly, Washington, DC and Kentucky are tied with the highest median alcoholic beer (ABV 6.25%), while New Mexico has the highest median bitterness (IBU 72).

```{r}
MergeBeerBreweries[which.max(MergeBeerBreweries$ABV),]
MergeBeerBreweries[which.max(MergeBeerBreweries$IBU),]
which.max(MedianABV)
MedianABV[c(8,18)]
which.max(MedianIBU)
MedianIBU[33]
```


##6. Summary statistics for the ABV variable.

```{r}
summary(MergeBeerBreweries$ABV)
```


##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

#### There appears to be a slight relationship between ABV and IBU (R^2 = 0.4461) 
#### (Note: slightly lower than our original analysis) 
#### The regression equation is (IBU = -34.66 + 1293.22*ABV)

#### For every 0.01 point increase in ABV, IBU increases by ~12.93 
#### (Note: slighlty higher than before)

```{r}
attach(MergeBeerBreweries)
library(ggplot2)
r <- ggplot(MergeBeerBreweries, na.rm = TRUE, aes (x = ABV, y = IBU)) + geom_point (color = "aquamarine2", size = 2) + geom_smooth (method = lm, color = '#2C3D50') 
r + ggtitle("Alcohol vs Bitterness in Beer") + xlab("Alcohol (ABV)") + ylab("Bitterness (IBU)")
relation <- lm(IBU~ABV)
print(summary(relation))
```

##Conclusion
###Our FrankenBeer has distinct qualities that make it appealing to many demographics of beer drinkers. Initial testing has been done in the Seattle, Washington and Boulder, Colorado markets with much success.
###West Virginia, New Mexico, and Florida represent the top three markets in terms of median bitterness by alcohol content.
###With its high octane ABV of 12% and its subtle IBU of 33, we believe our FrankenBeer will make its mark in these new markets.
