---
title: "Project 1"
author: "Allen Crane"
date: "June 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Set working directory and get data files

```{r}
setwd("C:\\Users\\allen\\Documents\\SMU Data Science\\MSDS6306 - Doing Data Science\\Lecture 5")
beers.df <- read.csv("Beers.csv", header=TRUE, sep = ",")
breweries.df <- read.csv("Breweries.csv", header=TRUE, sep = ",")
```


##1. How many breweries are present in each state?

```{r}
library(plyr)
brewery_count <- count(breweries.df, 'State')
brewery_count
```


##2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

#### First create a new variable "Brewery_id" (copy of "Brew_ID") that can merge both files
#### Merge data on Brewery_ID
#### Clean up column names

```{r}
breweries.df$Brewery_id <- breweries.df$Brew_ID
MergeBeerBreweries <- merge(x=beers.df, y=breweries.df, by="Brewery_id", all=TRUE)
names(MergeBeerBreweries) <- c("Brewery_ID", "Beer_Name", "Beer_ID", "ABV", "IBU", "Style", "Ounces", "Brew_ID", "Brewery_Name", "City", "State")
head(MergeBeerBreweries)
tail(MergeBeerBreweries)
```


##3. Report the number of NA's in each column.

```{r}
summary(MergeBeerBreweries)
```


##4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.

#### First create a median function so we can group by State
#### Median of ABV by State
#### Plot bar chart of ABV by State
#### Median of IBU by State
#### Plot bar chart of IBU by State


```{r}
MedianBeerStat <- function(x) {
  c(median = median(x[which(!is.na(x))]))
}
tapply(MergeBeerBreweries$ABV, MergeBeerBreweries$State, MedianBeerStat)
MedianABV <- tapply(MergeBeerBreweries$ABV, MergeBeerBreweries$State, MedianBeerStat)
barplot(MedianABV, main="Median Alcohol By Volume (ABV) by State", xlab="State", ylab="Median ABV", col="darkblue")
tapply(MergeBeerBreweries$IBU, MergeBeerBreweries$State, MedianBeerStat)
MedianIBU <- tapply(MergeBeerBreweries$IBU, MergeBeerBreweries$State, MedianBeerStat)
barplot(MedianIBU, main="Median International Bitterness Units (IBU) by State", xlab="State", ylab="Median IBU", col="darkblue")
```


##5. Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

#### Colorado has the maximum alcoholic (ABV) beer "Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale", in terms of individual beers, while Oregon has the most bitter beer "Bitter Bitch Imperial IPA", in terms of individual beers.
#### Using the median ABV and IBU data respectivly, Washington, DC and Kentucky are tied with the highest median alcoholic beer (ABV 0.0625), while Maine has the highest median bitterness (IBU 61).

```{r}
MergeBeerBreweries[which.max(MergeBeerBreweries$ABV),]
MergeBeerBreweries[which.max(MergeBeerBreweries$IBU),]
which.max(MedianABV)
MedianABV[c(8,18)]
which.max(MedianIBU)
MedianIBU[22]
```


##6. Summary statistics for the ABV variable.

```{r}
summary(MergeBeerBreweries$ABV)
```


##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

#### There appears to be a slight relationship between ABV and IBU (R^2 = 0.4493). 
#### The regression equation is (IBU = -34.1 + 1282*ABV)

#### For every 0.01 point increase in ABV, IBU increases by ~12.82.

```{r}
attach(MergeBeerBreweries)
plot(ABV, IBU, main="Alcohol vs Bitterness in Beer", xlab="Alcohol (ABV)", ylab="Bitterness (IBU)")
abline(lm(IBU~ABV), col="red") # regression line (IBU~ABV)
relation <- lm(IBU~ABV)
print(summary(relation))
```